### 2025-04-08 Paper.tex
final submission paper

    \vfill
    {\large\today\par}
\end{titlepage}

@@

	The singing voice conversion and the singing voice synthesis approach both come with advantages and difficulties, on which our model choice is based. SVC is a manipulation of the source audio and therefore often outperforms SVS models in preserving the singers expressiveness and naturalness, whereas SVS grant the user full control over the model and much more flexibility regarding the input songs (because it’s not reliant on existing performances). More flexibility increases the data and time consumption of synthesis models during training, which leads to higher computational costs compared with conversion models.

	Another significant factor was the public data availability for our project. We have plenty of training data for our target voice (Chester Bennington), including 7 studio albums, 3 live albums and even more live performances - but we encounter a sparse data situation for Emily’s voice, where only a single album is published and even less audios are available, where both perform the same song.

@@

	Apart from different approaches to enhance the target voice naturalness, low resource availability is a prevailing challenge in singing voice synthesis. Numerous models were designed to be more data efficient, requiring only small portions of training data, compared to the conventional models. The Sinsy \cite{Cho2021} model produces proper results with 1 hour of conditioned training data, partly due to a partition of the models architecture into further submodules. Zero shot models (for example StyleSinger and YourTTS) request merely minutes of training data to transfer voices in different languages or styles. \cite{Casanova2021},\cite{Zhang2024}

@@

	%==============================================
	%================ Preprocessing ================
	%==============================================
	\section{Preprocessing}

@@

	\subsubsection{Vocal Extraction}
 Music source separation consists in isolating single sound sources from a mixed audio signal. When the vocal track is extracted from a mix, this practice is usually called vocal extraction. \cite{Vincent2006} categorize source separation into techniques that rely on statistical assumptions (e.g., Independent Component Analysis) and the recent  models that rely on deep learning approaches. In the case of vocal extraction, the methods employed make use of time-frequency representations such as spectrogram masks or raw waveform modeling, with the aim to isolate the singer’s voice while minimizing artifacts and instrumental interferences \cite{Huang2014} Nevertheless, contemporary deep learning methods outperform traditional statistical approaches precisely because they do not operate under the assumptions of independence or basic spectral models. However, this assumption almost never holds true in real-world music recordings, as there is no independence between sources. Conversely, deep learning methods capture complex, nonlinear representations of music directly from data \cite{Uhlich2017}. By learning higher-level features (timbre and harmonic relationships) they are better able to adapt even to recordings, handling as a result more complex scenarios and producing cleaner vocal isolations. The separation process is considerably more robust than with statistical models that tend to falter when sources share similar frequency bands or show correlated dynamics \cite{Vincent2006}.

	The deep neural network approach used for extracting vocals or instruments from a mix can be broken down in two main strategies. The first one is the spectrogram-based separation which converts the audio waveform into a time-frequency representation (usually through the Short-Time Fourier Transform) and subsequently trains a neural network to predict masks or direct magnitudes for each source \cite{Stoeter202} Thus the aim of the spectrogram-based approach is to minimize cross-talk between instruments in the final extracted signal by leveraging both magnitude and phase information in each frequency bin. As a result, each audio frame is split into small frequency bins, each corresponding to a discrete range of frequencies. On the other hand, the second strategy is based on waveform-based separation in the time domain. For instance, Demucs introduced an encoder-decoder architecture operating on raw waveforms that preserves amplitude envelopes more effectively than frequency-domain approaches (Défossez et al., 2019). Yet, while spectrogram-based methods have proven more suitable for certain genres and recordings, time-domain networks offer a superior performance when dealing with fast transients (e.g., drum hits) or complex stereo imaging.

@@

	\subsubsection{Segmentation}
	After the vocal extraction, the next step was to segment the data. Segmenting the audios into smaller chunks is important as it can significantly enhance the model’s performance and the reliability of the forced alignment that followed this step \cite{Rybach2009}. The quality of the alignment heavily depends on the segmentation, as a wrong segmentation will lead to a wrong alignment. The effectiveness of our model is, therefore, closely tied to the goodness of the audio segmentation.

@@

	Following an initial phase of trial and error, we implemented a custom script utilizing whisper-timestamped to segment our audio data. The script first verifies the file format, ensuring that only audio files (with the ending wav or mp3) are processed and that only those files whose names conformed to our naming convention (i.e., filenames containing either six or nine digits) were included. Once the appropriate files were selected, we used whisper-timestamped to generate segment-level transcriptions along with time-aligned boundaries. Most of the segments were fine, but some segments were extremely short (i.e., less than two seconds). To avoid these short segments and to enhance the usability of the resulting segments for annotation purposes, we introduced a minimum segment duration of four seconds. Thus, segments falling below this threshold were automatically concatenated with the subsequent segment to ensure that each snippet was sufficiently long to carry meaningful context for the automatic labeling. We also had some issues with hard cuts, where word endings were cut off too strictly. To mitigate this problem, we introduced a small buffer between consecutive segments (0.1s). This slight overlap helped preserve the natural flow of speech and ensured smoother transitions, which improved both the accuracy of our transcriptions and the overall quality of the audio snippets.

@@

	To improve alignment quality, we explored alternative tools and came across SOFA (Singing-Oriented Forced Aligner), which functions similarly to the Montreal Forced Aligner (MFA) but is specifically designed for singing voice. According to its Github page, SOFA offers simpler installation, better performance, and faster inference speed compared to MFA. After testing it on a few examples, we observed slightly improved results \cite{Greenleaf2001}; however, the alignments were still not consistently convincing. We continued our search and discovered LabelMakr, “a GUI tool to help users easily generate SVS phoneme-level labels” \cite{spicytigermeat} for singing voice synthesis (SVS), particularly intended for use with DiffSinger. While initial tests with a small sample also fell short of expectations, we observed that LabelMakr's performance significantly improved when it was applied to our full dataset, such that we ended up with some meaningful alignments that we used for our training.

@@

	\begin{figure}[htbp]
		\centering
		\includegraphics[width=0.5\textwidth]{graphics/manual_alignment.png}
		\caption{manual alignment}
		\label{fig:bild3}
	\end{figure}

	\subsubsection{Data Conversion with nnsvs-db-converter}
	We used the nnsvs-db-converter, a Python tool for converting vocal databases with HTS mono labels into DiffSinger format to make them compatible with variance models for predicting pitch and timing \cite{nnsvsdbconverter}. It handles sample segmentation by ensuring that each sample has a maximum length and an acceptable number of pauses between segments, and converts silent phonemes (e.g., sil, pau) to standardized representations (SP and AP for breaths). The script requires minimal external dependencies, making it easily accessible. In addition, it supports language definitions that specify vowels and liquids for better segmentation and timing handling, and it can estimate MIDI pitch and timing information for the use of variance models.

@@

	Initially, we attempted to map each song snippet to the category it was most similar to. However, this yielded inconsistent and often unreliable classifications, such that the categories had a large overlap and seemed to be meaningless. We then tried to identify dominant acoustic features that could meaningfully distinguish between categories. We found that fundamental frequency was a strong discriminator between screaming and clear singing. After analyzing our reference examples, we set a threshold of 230 Hz to filter out heavy screams and background noise—this removed many problematic segments while retaining the majority of clean vocals. The remaining screams in the clean voice examples were retained as they were less problematic than the really heavy ones and the background noises.

	To distinguish between rap and singing, we found that syllable rate is a suitable discriminator. Since our target singer, Chester Bennington, does not perform the rap sections, we applied a threshold of six syllables per second to exclude the fast-paced vocal segments.

	For segments with multiple overlapping voices, we explored various characteristics and thresholds, but none of them seemed to be a reliable discriminator. Therefore, we returned to the similarity-based approach—this time using only a single, well-chosen example per category to guide the classification, which surprisingly seemed to work better than with the bigger example set. While the categories were still imperfect, this method provided the most usable approximation.

@@

	%==============================================
	%================ Training =====================
	%==============================================
	\section{Training}

@@

	\begin{itemize}
		\item \textbf{Pitch Shifting}: We wrote a script to adjust the pitch of each audio file by a semitone both upwards and downwards, creating two additional variations per original file. This technique helps the model become invariant to pitch differences as the pitch changes heavily between Chester’s different singing modes (e.g., compare harmonic singing versus screaming). We also tried to make more drastic pitch shifting, however, then the voice seemed to be too different from the original, so we did not include it in our model training.

		\item \textbf{Loudness Adjustment}: The audio files’ loudness was modified by increasing and decreasing the volume by 10\%. This approach aids the model in handling variations in speaking volume and recording conditions.

@@

	The LST cluster of the Saarland University uses HTCondor to schedule jobs and distributes them to free GPUs in incoming order. Users request the necessary computational cost for a job via submit files and transfer the file to be executed (must be a bash script). The dependencies are managed inside the bash script with an anaconda environment or in the submit file using a docker. For this purpose, we created an anaconda environment running on python version 3.8.20 and installed the necessary requirements following the DiffSinger installation guide. \cite{openvpi-diffsinger}

	Though HTCondor provides the possibility to distribute the executed job among several GPUs to decrease the running time, we restricted our training to 1 GPU (less potential to run into errors) and requested 16GB of GPU memory. After running for a maximum of 1 day, jobs get automatically kicked from the server by default. The image shows an example submit file that we used in our project.

@@

	The main architecture of the DiffSinger model is trained to predict mel-spectrograms out of the input data, but that is not exactly what we challenged ourselves. To generate audio files from the mel-spectrograms a vocoder is necessary. For a start we stuck to the HifiGAN vocoder recommended in the openvpi DiffSinger support, which we ultimately fine-tuned on Chester’s voice (training see below in the section vocoder).

	The repository offers several built-in models operating as pitch extractor and harmonic-noise separator as well, recommending RMVPE for extracting pitch and Vocal Remover (VR) to separate the harmonic and aperiodic parts of the singer's voice. For the shallow diffusion configuration again we were oriented strongly towards the given guide, copying the preset parameter values.

@@

	\subsection{Used files}
	The following 14 segments were used in the evaluation:
	\begin{itemize}
		\item Segment 22 from album 1, song 4
		\item Segment 3 from album 1, song 5
		\item Segment 1 from album 1, song 8
		\item Segment 22 from album 1, song 8
		\item Segment 5 from album 2, song 4
		\item Segment 7 from album 2, song 6
		\item Segment 17 from album 2, song 6
		\item Segment 1 from album 2, song 9
		\item Segment 7 from album 2, song 9
		\item Segment 3 from album 3, song 5
		\item Segment 13 from album 3, song 5
		\item Segment 1 from album 3, song 6
		\item Segment 10 from album 3. song 6
		\item Segment 8 from album 5, song 3
	\end{itemize}

	\subsection{Criteria}

@@

	The variance+vocoder model outperformed all other models and reached results comparable to the ground truth, indicating outputs with relatively high audio quality. At some distance follows the result of the manual+clean model, achieving a slightly worse but still comparably high predicted MOS score. What is striking is that in particular models trained on smaller datasets (independent of audio quality) seem to receive lower MOS prediction values. We suspect this trend to be provoked by noise in the output audios.

	\subsubsection{Histogram Comparison}

@@

	While keeping in mind that the Intersection value might possibly be less informative than other comparison measurements, the variance+vocoder and large model again achieve the highest results, with the most overlapping of bin values. Furthermore, models trained or fine-tuned on the clean dataset perform the worst with this metric. Potentially, the “high quality” data is shifting or skewing the results in either direction, which preserves general distribution properties but changes the bin overlap.

	The Chi-Square distance remains the only measurement, where the variance+vocoder model is beaten by other models - with the small and large model outperforming the rest, both reaching values below 300000. However, no enlightening tendencies or correlations between model hyperparameters and Chi-square distance are observable, therefore the values appear to be rather meaningless for the evaluation.
	The variance+vocoder model scores the smallest Bhattacharyya distance with the value 0.16, once more outperforming all other tested models. In general models with the fine-tuned vocoder reached comparatively good results, where further research is necessary to investigate potential reasons for this effect.

@@

	Despite extensive experimentation, the overall model performances remain relatively close. While some vocal segments are handled well by most models, others result in distorted, rough-sounding outputs or noticeable background noise. Common issues that we encounted during inference include misaligned or incorrect phonemes and inconsistent transitions. To quantify the model performances, we used five evaluation methods: predicted Mean Opinion Score (MOS) via SingMOS, mel-spectrogram histogram comparisons, objective MOS using PESQ, F0 Root Mean Square Error (F0 RMSE), and Speech-to-Noise Ratio (SNR). The variance+vocoder model revealed the best MOS scores and a strong spectral similarity to real vocals. The manual+sorted clean model achieved the best results for PESQ. Model 8 (manual+sorted+augmented clean) achieved the best results in the F0 RMSE evaluation and the large model achieved the best results for the SNR evaluation. This shows that the models have different strengths and weaknesses and in the end achieved quite similar results. Even the top-performing models could not fully replicate the expressive detail and clarity of the original vocals, often struggling with distorted phonemes, rough transitions, or alignment errors.

>>>>

    \vfill     
    {\large April 8, 2025\par}
\end{titlepage}

@@

	The singing voice conversion and the singing voice synthesis approach both come with advantages and difficulties, on which our model choice is based. SVC is a manipulation of the source audio and therefore often outperforms SVS models in preserving the singers expressiveness and naturalness, whereas SVS grant the user full control over the model and much more flexibility regarding the input songs (because it’s not reliant on existing performances). More flexibility increases the data and time consumption of synthesis models during training, which leads to higher computational costs compared with conversion models. 
	\newpage
	Another significant factor was the public data availability for our project. We have plenty of training data for our target voice (Chester Bennington), including 7 studio albums, 3 live albums and even more live performances - but we encounter a sparse data situation for Emily’s voice, where only a single album is published and even less audios are available, where both perform the same song.

@@

	Apart from different approaches to enhance the target voice naturalness, low resource availability is a prevailing challenge in singing voice synthesis. Numerous models were designed to be more data efficient, requiring only small portions of training data, compared to the conventional models. The Sinsy \cite{Cho2021} model produces proper results with 1 hour of conditioned training data, partly due to a partition of the models architecture into further submodules. Zero shot models (for example StyleSinger and YourTTS) request merely minutes of training data to transfer voices in different languages or styles. \cite{Casanova2021},\cite{Zhang2024}
	\newpage

@@

	\newpage
	%==============================================
	%================ Preprocessing ================
	%==============================================
	\section{Preprocessing}

@@

	\subsubsection{Vocal Extraction}
 Music source separation consists in isolating single sound sources from a mixed audio signal. When the vocal track is extracted from a mix, this practice is usually called vocal extraction. \cite{Vincent2006} categorize source separation into techniques that rely on statistical assumptions (e.g., Independent Component Analysis) and the recent  models that rely on deep learning approaches. In the case of vocal extraction, the methods employed make use of time-frequency representations such as spectrogram masks or raw waveform modeling, with the aim to isolate the singer’s voice while minimizing artifacts and instrumental interferences \cite{Huang2014} Nevertheless, contemporary deep learning methods outperform traditional statistical approaches precisely because they do not operate under the assumptions of independence or basic spectral models. However, this assumption almost never holds true in real-world music recordings, as there is no independence between sources. Conversely, deep learning methods capture complex, nonlinear representations of music directly from data \cite{Uhlich2017}. By learning higher-level features (timbre and harmonic relationships) they are better able to adapt even to recordings, handling as a result more complex scenarios and producing cleaner vocal isolations. 
    \newpage
The separation process is considerably more robust than with statistical models that tend to falter when sources share similar frequency bands or show correlated dynamics \cite{Vincent2006}.

	The deep neural network approach used for extracting vocals or instruments from a mix can be broken down in two main strategies. The first one is the spectrogram-based separation which converts the audio waveform into a time-frequency representation (usually through the Short-Time Fourier Transform) and subsequently trains a neural network to predict masks or direct magnitudes for each source \cite{Stoeter202} Thus the aim of the spectrogram-based approach is to minimize cross-talk between instruments in the final extracted signal by leveraging both magnitude and phase information in each frequency bin. As a result, each audio frame is split into small frequency bins, each corresponding to a discrete range of frequencies. On the other hand, the second strategy is based on waveform-based separation in the time domain. For instance, Demucs introduced an encoder-decoder architecture operating on raw waveforms that preserves amplitude envelopes more effectively than frequency-domain approaches \cite{Defossez2019}. Yet, while spectrogram-based methods have proven more suitable for certain genres and recordings, time-domain networks offer a superior performance when dealing with fast transients (e.g., drum hits) or complex stereo imaging.

@@

 	\newpage
	\subsubsection{Segmentation}
	After the vocal extraction, the next step was to segment the data. Segmenting the audios into smaller chunks is important as it can significantly enhance the model’s performance and the reliability of the forced alignment that followed this step \cite{Rybach2009}. The quality of the alignment heavily depends on the segmentation, as a wrong segmentation will lead to a wrong alignment. The effectiveness of our model is, therefore, closely tied to the goodness of the audio segmentation.

@@

	Following an initial phase of trial and error, we implemented a custom script utilizing whisper-timestamped to segment our audio data. The script first verifies the file format, ensuring that only audio files (with the ending wav or mp3) are processed and that only those files whose names conformed to our naming convention (i.e., filenames containing either six or nine digits) were included. Once the appropriate files were selected, we used whisper-timestamped to generate segment-level transcriptions along with time-aligned boundaries. Most of the segments were fine, but some segments were extremely short (i.e., less than two seconds). 
	\newpage
To avoid these short segments and to enhance the usability of the resulting segments for annotation purposes, we introduced a minimum segment duration of four seconds. Thus, segments falling below this threshold were automatically concatenated with the subsequent segment to ensure that each snippet was sufficiently long to carry meaningful context for the automatic labeling. We also had some issues with hard cuts, where word endings were cut off too strictly. To mitigate this problem, we introduced a small buffer between consecutive segments (0.1s). This slight overlap helped preserve the natural flow of speech and ensured smoother transitions, which improved both the accuracy of our transcriptions and the overall quality of the audio snippets.

@@

	To improve alignment quality, we explored alternative tools and came across SOFA (Singing-Oriented Forced Aligner), which functions similarly to the Montreal Forced Aligner (MFA) but is specifically designed for singing voice. According to its Github page, SOFA offers simpler installation, better performance, and faster inference speed compared to MFA. After testing it on a few examples, we observed slightly improved results \cite{Greenleaf2001}; however, the alignments were still not consistently convincing. 
	\newpage
We continued our search and discovered LabelMakr, “a GUI tool to help users easily generate SVS phoneme-level labels” \cite{spicytigermeat} for singing voice synthesis (SVS), particularly intended for use with DiffSinger. While initial tests with a small sample also fell short of expectations, we observed that LabelMakr's performance significantly improved when it was applied to our full dataset, such that we ended up with some meaningful alignments that we used for our training.

@@

	\begin{figure}[htbp]
		\centering
		\includegraphics[width=0.5\textwidth]{graphics/manual_alignment.png}
		\caption{manual alignment}
		\label{fig:bild3}
	\end{figure}
	\newpage
	\subsubsection{Data Conversion with nnsvs-db-converter}
	We used the nnsvs-db-converter, a Python tool for converting vocal databases with HTS mono labels into DiffSinger format to make them compatible with variance models for predicting pitch and timing \cite{nnsvsdbconverter}. It handles sample segmentation by ensuring that each sample has a maximum length and an acceptable number of pauses between segments, and converts silent phonemes (e.g., sil, pau) to standardized representations (SP and AP for breaths). The script requires minimal external dependencies, making it easily accessible. In addition, it supports language definitions that specify vowels and liquids for better segmentation and timing handling, and it can estimate MIDI pitch and timing information for the use of variance models.

@@

	Initially, we attempted to map each song snippet to the category it was most similar to. However, this yielded inconsistent and often unreliable classifications, such that the categories had a large overlap and seemed to be meaningless. We then tried to identify dominant acoustic features that could meaningfully distinguish between categories. We found that fundamental frequency was a strong discriminator between screaming and clear singing. After analyzing our reference examples, we set a threshold of 230 Hz to filter out heavy screams and background noise—this removed many problematic segments while retaining the majority of clean vocals. The remaining screams in the clean voice examples were retained as they were less problematic than the really heavy ones and the background noises.
	\newpage
	To distinguish between rap and singing, we found that syllable rate is a suitable discriminator. Since our target singer, Chester Bennington, does not perform the rap sections, we applied a threshold of six syllables per second to exclude the fast-paced vocal segments.
	For segments with multiple overlapping voices, we explored various characteristics and thresholds, but none of them seemed to be a reliable discriminator. Therefore, we returned to the similarity-based approach—this time using only a single, well-chosen example per category to guide the classification, which surprisingly seemed to work better than with the bigger example set. While the categories were still imperfect, this method provided the most usable approximation.

@@

	\newpage
	%==============================================
	%================ Training =====================
	%==============================================
	\section{Training}

@@

	\begin{itemize}
		\item \textbf{Pitch Shifting}: We wrote a script to adjust the pitch of each audio file by a semitone both upwards and downwards, creating two additional variations per original file. This technique helps the model become invariant to pitch differences as the pitch changes heavily between Chester’s different singing modes (e.g., compare harmonic singing versus screaming). We also tried to make more drastic pitch shifting, however, then the voice seemed to be too different from the original, so we did not include it in our model training.
		\newpage
		\item \textbf{Loudness Adjustment}: The audio files’ loudness was modified by increasing and decreasing the volume by 10\%. This approach aids the model in handling variations in speaking volume and recording conditions.

@@

	The LST cluster of the Saarland University uses HTCondor to schedule jobs and distributes them to free GPUs in incoming order. Users request the necessary computational cost for a job via submit files and transfer the file to be executed (must be a bash script). The dependencies are managed inside the bash script with an anaconda environment or in the submit file using a docker. For this purpose, we created an anaconda environment running on python version 3.8.20 and installed the necessary requirements following the DiffSinger installation guide. \cite{openvpi-diffsinger}
	\newpage
	Though HTCondor provides the possibility to distribute the executed job among several GPUs to decrease the running time, we restricted our training to 1 GPU (less potential to run into errors) and requested 16GB of GPU memory. After running for a maximum of 1 day, jobs get automatically kicked from the server by default. The image shows an example submit file that we used in our project.

@@

	The main architecture of the DiffSinger model is trained to predict mel-spectrograms out of the input data, but that is not exactly what we challenged ourselves. To generate audio files from the mel-spectrograms a vocoder is necessary. For a start we stuck to the HifiGAN vocoder recommended in the openvpi DiffSinger support, which we ultimately fine-tuned on Chester’s voice (training see below in the section vocoder). 
	\newpage
	The repository offers several built-in models operating as pitch extractor and harmonic-noise separator as well, recommending RMVPE for extracting pitch and Vocal Remover (VR) to separate the harmonic and aperiodic parts of the singer's voice. For the shallow diffusion configuration again we were oriented strongly towards the given guide, copying the preset parameter values.

@@

	\subsection{Used files}
	The following 14 segments were used in the evaluation:
	\begin{itemize}
		\item Segment 22 from album 1, song 4
		\item Segment 3 from album 1, song 5
		\item Segment 1 from album 1, song 8
		\item Segment 22 from album 1, song 8
		\item Segment 5 from album 2, song 4
		\item Segment 7 from album 2, song 6
		\item Segment 17 from album 2, song 6
		\item Segment 1 from album 2, song 9
		\item Segment 7 from album 2, song 9
		\item Segment 3 from album 3, song 5
		\item Segment 13 from album 3, song 5
		\item Segment 1 from album 3, song 6
		\item Segment 10 from album 3. song 6
		\item Segment 8 from album 5, song 3
	\end{itemize}
	\newpage
	\subsection{Criteria}

@@

	The variance+vocoder model outperformed all other models and reached results comparable to the ground truth, indicating outputs with relatively high audio quality. At some distance follows the result of the manual+clean model, achieving a slightly worse but still comparably high predicted MOS score. What is striking is that in particular models trained on smaller datasets (independent of audio quality) seem to receive lower MOS prediction values. We suspect this trend to be provoked by noise in the output audios.
	\newpage
	\subsubsection{Histogram Comparison}

@@

	While keeping in mind that the Intersection value might possibly be less informative than other comparison measurements, the variance+vocoder and large model again achieve the highest results, with the most overlapping of bin values. Furthermore, models trained or fine-tuned on the clean dataset perform the worst with this metric. Potentially, the “high quality” data is shifting or skewing the results in either direction, which preserves general distribution properties but changes the bin overlap.
	\newpage
	The Chi-Square distance remains the only measurement, where the variance+vocoder model is beaten by other models - with the small and large model outperforming the rest, both reaching values below 300000. However, no enlightening tendencies or correlations between model hyperparameters and Chi-square distance are observable, therefore the values appear to be rather meaningless for the evaluation.
	The variance+vocoder model scores the smallest Bhattacharyya distance with the value 0.16, once more outperforming all other tested models. In general models with the fine-tuned vocoder reached comparatively good results, where further research is necessary to investigate potential reasons for this effect.

@@

	Despite extensive experimentation, the overall model performances remain relatively close. While some vocal segments are handled well by most models, others result in distorted, rough-sounding outputs or noticeable background noise. Common issues that we encounted during inference include misaligned or incorrect phonemes and inconsistent transitions. To quantify the model performances, we used five evaluation methods: predicted Mean Opinion Score (MOS) via SingMOS, mel-spectrogram histogram comparisons, objective MOS using PESQ, F0 Root Mean Square Error (F0 RMSE), and Speech-to-Noise Ratio (SNR). The variance+vocoder model revealed the best MOS scores and a strong spectral similarity to real vocals. The manual+sorted clean model achieved the best results for PESQ. Model 8 (manual+sorted+augmented clean) achieved the best results in the F0 RMSE evaluation and the large model achieved the best results for the SNR evaluation. 
	\newpage
This shows that the models have different strengths and weaknesses and in the end achieved quite similar results. Even the top-performing models could not fully replicate the expressive detail and clarity of the original vocals, often struggling with distorted phonemes, rough transitions, or alignment errors.
